A Graph Theory Based Approach to MDAO Problem Formulation

### 2013-04-10 graph_syntax.tex
cross referencing

\section{Requirements for New Graph Syntax}
The ultimate goal of the new graph--based syntax presented here is to be able to 
fully describe the general structure of an MDAO problem independently of any solution information, 
while still being able to accomodate the more specific case when a solution 
strategy is applied. In order to achieve that goal 
the graph needs to accomodate a number of constructs of MDAO problems: 
\begin{itemize}
    \item Analysis tools and connections between them
    \item Design variables, objectives, and constraints
    \item Local and global properties
    \item Coupling between analyses
    \item Multi-fidelity analyses
\end{itemize}

@@

%the relevant MDAO constructs from the given grah primatives. Next we show how to construct 
%the MCG, FPG, and SPF graphs from those constructs and demonstrate methods to transition between 
%them.

\section{Syntax Definition}
In this section we present the necessary graph theory fundamentals to 
construct the graphs discussed above. 
The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
For a \emph{directed graph*} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
of a set of sets. Each ordered pair represents an edge starting at the node 
indicated by the first entry and directed to the node indicated by the second 
entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
\begin{figure}[htb!]
	\begin{center}
	\includegraphics[width=1.5in]{images/example_directed_graph}
	\end{center}
	\vspace{-20pt}
\caption{Example directed graph.}
\label{f:example directed graph}
\end{figure}
As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
\begin{IEEEeqnarray*}{rCl}
V & = & \{1,2,3,4,5\}, \\
E & = & \big\{(1,2),(3,2),(1,3),(3,4)\big\}.
\end{IEEEeqnarray*}

@@

\subsection{Node and Edge Types}
We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
There are three node types:  
\begin{description}
\item[variable: ] represents scalar or array data
\item[model:] responsible for mapping inputs to outputs
\item[driver:] control logic capable of managing iteration
\end{description}

@@

\subsection{Rules for Nodes and Edges}
There are specific rules for the usage of these nodes and edges.
The driver node and the implicit edge are only allowed to be present in PSGs. All 
other node and edge types can be present in any of the three graphs, subject to these restrictions: 
\begin{enumerate}
\item A model node can only have one edge directed to or from another model node.
\item A model node can only have fixed connection edges directed in or out.
\item A model node must have at least one edge directed in and at least one edge 
  directed out.
\item If a variable node has an outgoing edge to a model node then it may not have 
  any other outgoing edges.
\end{enumerate}

@@

  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
    \begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
	This represents an overload of information such that it is unclear which source to use.
\end{description} 
An algorithm for detecting and managing these violations if proposed in Sec.(()).

\section{Graph Representation of MDAO Constructs}
%As discussed in Sec.((intro)), the digraph must be able to describe the individual aspects of an MDAO problem.
\subsection{Analysis Blocks and Connections}
Analysis tools take in a number input variables and then perform some work to calculate 
the values for their respective outputs. In an MDAO graph, this process is 
represented by a group of nodes and edges called an \emph{analysis block}, 
shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
node represents a single input or output and is connected 
to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
the analysis block contains two model nodes, with a single edge connecting them. 
This edge represents the necessary calculations to map given inputs 
to proper output values. This edge can be given a weight to represent the computational cost of the analysis block. However, it is also allowed to skip 
this edge and have both incomming and outgoing edges to variables through a single model 
node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
the edges within them are fixed, they become fixed sub-graphs within a graph.

\begin{figure}[htb!]
    \begin{center}
    \includegraphics[width=4in]{images/analysis_block}
    \end{center}
    \vspace{-10pt}
\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
\label{f:analysis block}
\end{figure}

@@

\subsection{Design Variables}
((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
Design variables require external decisions 
to be made about a given system to identify them. Consider 
a notional problem given by the graph in Fig.~\ref{f:designvars}. There are three input variables 
which could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
 $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
have no incomming edges and could each be selected as design variables. 
\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/design_vars_graph}
  \end{center}
  \caption{Notional graph with two potential design variables \label{f:designvars}}
\end{figure}

@@

\subsection{Objectives, and Constraints}
Objectives and constraints are also MDAO constructs that need to be identified by engineers 
as part of the design process. In the case of objective functions single output values 
could be selected, but commonly multiple values are assembeled together via simple 
algebraic expression to form a objective function. For example, the problem depicted in 
Fig.~\ref{f:designvars} could have an objective function of $z^2$ or $2y + z$. Constraints are 
always given in the form of either an inequality of an equality, e.g. 
$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
algebraic expression that can be evaluated to produce an output.

These operations, though very simple, are effectively the same kind of task 
performed by an analysis block. A constraint or objective function can therfore 
be represented as another analysis block on the graph, with it's own input and 
output variable nodes. Although fundamentally no different than an analysis block, 
for clarity and convience it is useful to distinguish between regular analysis 
blocks and those that arise from the addition of objectives or constraints to 
the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
shows a notional problem graph augmented with expression blocks and appropriate 
connection edges.

\begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=.6\textwidth]{images/obj_const_graph}
  \end{center}
  \caption{Notional graph with and objective and constraint nodes \label{f:obj-cons}}
\end{figure}

\subsection{Local vs Global Nodes}

  A simple definition of a global node is any node that directly involves information 
  from more than one analysis block. Likewise, a local node is any node 
  that involves information from only a single analysis block. As stated before, variable 
  nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
  and similarly model nodes are also local. Expression and driver nodes
  are not inherently local or global though. For these, their locality is determined 
  by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
  where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
  Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
  on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint.

  \begin{figure}
      \begin{center}
      %\includegraphics[height=.25\textheight]{XDSM/simple}
      [[insert graph here]]
      \caption{A simple graph with local and global nodes \label{f:local_global}}
      \end{center}
  \end{figure}

@@

\subsection{Coupling Between Analyses}

  Coupling exists in a design problem when a set of two or more analyses each depend on the
  others outputs. Consider the notional problem from Sec.~\ref{sec:spg_vs_fpg} where two 
  analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
  is shown in Fig. \ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
  edge to $B.z$, indicating that there is a connection between the two. 
  Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
  a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
  to be iterated on until they converge.

  \begin{figure}
      \begin{center}
      %\includegraphics[height=.25\textheight]{}
      [INSERT GRAPH HERE]
      \caption{Graph of a notional problem with simple coupling \label{f:coupling}}
      \end{center}
  \end{figure}

@@

  Larger problems can contain more complex cycles in their FPG, indicating more 
  complex coupling between analyses. A cycle can contain more than 
  just two analysis blocks. Multiple independent cycles can exist, indicating 
  independent coupling relationships. Cycles can also overlap, where the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. From the perspective of the graph, the problem reduces down to 
  an ordering challenge. For each cycle in the graph, an analysis block needs to be selected 
  to start the iteration from. Rogers' DEMAID tool uses a genetic algorithm to find an ordering 
  that minimizes the overall coupling of the system by separating independent 
  cycles in the graph \cite{rogers1996,rogers1996demaid}. In the present work, Sec.(()) will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.

\subsection{Multi--fidelity Problems}

  A multi--fidelity problem can be characterized by two or more different analyses 
  each calculating the same data. Mapping this type of multi--fidelity situation 
  onto a graph will yield two analysis blocks, each with their own output 
  variable nodes, having outgoing edges to a third analysis block that needs the 
  data as input. Figure \ref{f:collision-example} shows a notional problem 
  that exhibits this structure, where the key feature is that variable node $y$ in 
  analysis block $C$ has two incoming edges.

  \begin{figure}
      \begin{center}
      %\includegraphics[height=.25\textheight]{}
      [INSERT GRAPH HERE]
      \caption{Graph of a notional multi-fidelity problem \label{f:collision-example}}
      \end{center}
  \end{figure}

@@

\section{Building MCG and FPG}

\subsection{Maximal Connectivity Graph}
  To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided. The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$. The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$. 
  We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
  Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
  \begin{IEEEeqnarray*}{rCl}
  V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
  E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
  \end{IEEEeqnarray*}
  The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

\subsection{Fundamental Problem Formulation Graph}
We now define the fundamental problem formulation graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
\begin{enumerate}
\item[(1)] $\displaystyle{I_F \subset I}$ such that the remaining conditions hold
\item[(2)] $\displaystyle{ \forall i \in I_F, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
\end{enumerate}
The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
Requirement (3) stipulates that only the global inputs that are being used should be included. 
Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper in-degree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

@@

\subsection{Obtaining the Fundamental Problem Graph}
In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all. Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. Then the problem is reduced to deciding which connection edges to remove.

>>>>

\section{Requirements for New Graph Syntax}
	\label{s:requirements}
	The ultimate goal of the new graph--based syntax presented here is to be able to 
	fully describe the general structure of an MDAO problem independently of any solution information, 
	while still being able to accomodate the more specific case when a solution 
	strategy is applied. In order to achieve that goal 
	the graph needs to accomodate a number of constructs of MDAO problems: 
	\begin{itemize}
		\item Analysis tools and connections between them
		\item Design variables, objectives, and constraints
		\item Local and global properties
		\item Coupling between analyses
		\item Multi-fidelity analyses
	\end{itemize}

@@

\section{Syntax Definition}
	\label{s:syntax definition}
	In this section we present the necessary graph theory fundamentals to 
	construct the graphs discussed above. 
	The notation used in this work is adapted from Diestel \cite{Diestel2010}. 
	A \emph{graph} is a pair $G = (V,E)$ of sets such that $E \subseteq V \times V$, 
	which means that the elements of $E$ are 2--element subsets of $V$. The set $V$ 
	contains the \emph{vertices} or \emph{nodes} and the set $E$ contains the \emph{edges}.
	For a \emph{directed graph*} (or \emph{digraph}) we construct $E$ as a set of ordered pairs instead 
	of a set of sets. Each ordered pair represents an edge starting at the node 
	indicated by the first entry and directed to the node indicated by the second 
	entry. Edge $e$ = $(x,y)$ may be referred to simply as $xy$. For node $v \in V$ 
	the edges directed out are given by $E(v)$ and the edges directed into $v$ are given 
	by $E^{-1}(v)$; $E(E(v))$ is denoted as $E^2(v)$, and likewise for additional levels. 
	If $E$ is not a one--to--one mapping, $E(v)$ may be the empty set, a single node, or a set of nodes.
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=1.5in]{images/example_directed_graph}
		\end{center}
		\vspace{-20pt}
	\caption{Example directed graph.}
	\label{f:example directed graph}
	\end{figure}
	As an example, for the directed graph shown in Fig.~\ref{f:example directed graph} we have
	\begin{IEEEeqnarray*}{rCl}
	V & = & \{1,2,3,4,5\}, \\
	E & = & \big\{(1,2),(3,2),(1,3),(3,4)\big\}.
	\end{IEEEeqnarray*}

@@

\subsection{Node and Edge Types}
	\label{ss:types}
	We now define different types of nodes and edges with specific properties to compose graphs used to represent an MDAO problem. 
	There are three node types:  
	\begin{description}
	\item[variable: ] represents scalar or array data
	\item[model:] responsible for mapping inputs to outputs
	\item[driver:] control logic capable of managing iteration
	\end{description}

@@

\subsection{Rules for Nodes and Edges}
	\label{ss:rules}
	There are specific rules for the usage of these nodes and edges.
	The driver node and the implicit edge are only allowed to be present in PSGs. All 
	other node and edge types can be present in any of the three graphs, subject to these restrictions: 
	\begin{enumerate}
	\item A model node can only have one edge directed to or from another model node.
	\item A model node can only have fixed connection edges directed in or out.
	\item A model node must have at least one edge directed in and at least one edge 
	  directed out.
	\item If a variable node has an outgoing edge to a model node then it may not have 
	  any other outgoing edges.
	\end{enumerate}

@@

	  \item[collision: ] The number of incomming edges is greater than $ \txt{deg}_u^-(v)$. 
		\begin{equation} \txt{deg}^-(v) > \txt{deg}_u^-(v) \end{equation}
		This represents an overload of information such that it is unclear which source to use.
	\end{description} 
	An algorithm for detecting and managing these violations if proposed in Sec.~\ref{s:building graphs}.\ref{ss:obtaining FPG}.

\section{Graph Representation of MDAO Constructs}
	\label{s:graph representation}
	%As discussed in Sec.~\ref{s:intro}, the digraph must be able to describe the individual aspects of an MDAO problem.
\subsection{Analysis Blocks and Connections}
	\label{ss:analysis blocks and connections}
	Analysis tools take in a number input variables and then perform some work to calculate 
	the values for their respective outputs. In an MDAO graph, this process is 
	represented by a group of nodes and edges called an \emph{analysis block}, 
	shown in Fig. \ref{f:analysis block}. Within an analysis block each variable 
	node represents a single input or output and is connected 
	to a single model node via a fixed edge. Note that in Fig. \ref{f:analysis block}, 
	the analysis block contains two model nodes, with a single edge connecting them. 
	This edge represents the necessary calculations to map given inputs 
	to proper output values. This edge can be given a weight to represent the computational cost of the analysis block. However, it is also allowed to skip 
	this edge and have both incomming and outgoing edges to variables through a single model 
	node for a given analysis block. Although analysis blocks are comprised of multiple nodes and edges, since all 
	the edges within them are fixed, they become fixed sub-graphs within a graph.
	\begin{figure}[htb!]
		\begin{center}
		\includegraphics[width=4in]{images/analysis_block}
		\end{center}
		\vspace{-10pt}
	\caption{Example analysis block. The each node type and edge type is labeled in italics and annotated parenthetically.}
	\label{f:analysis block}
	\end{figure}

@@

\subsection{Design Variables}
	\label{ss:design variables}
	((DP:I'm thinking we might also need the upper indegree limit to be zero as well, because setting only the lower indegree limit makes the node a design variable if it didn't have any incoming edges. It's seems possible that in the MCG a node has incoming edges but the user wants it to be a design variable. I can make the change if you agree.))
	Design variables require external decisions 
	to be made about a given system to identify them. Consider 
	a notional problem given by the graph in Fig.~\ref{f:designvars}. There are three input variables 
	which could potentially become design variables: $w$, $x$, and $y$. Because the variable $x$ is connected from $A$ to $B$,
	 $x$ can't actually be a design variable. On the other hand, the inputs $w$ and $y$
	have no incomming edges and could each be selected as design variables. 
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/design_vars_graph}
	  \end{center}
	  \caption{Notional graph with two potential design variables \label{f:designvars}}
	\end{figure}

@@

\subsection{Objectives, and Constraints}
	\label{ss:objectives and constraints}
	Objectives and constraints are also MDAO constructs that need to be identified by engineers 
	as part of the design process. In the case of objective functions single output values 
	could be selected, but commonly multiple values are assembeled together via simple 
	algebraic expression to form a objective function. For example, the problem depicted in 
	Fig.~\ref{f:designvars} could have an objective function of $z^2$ or $2y + z$. Constraints are 
	always given in the form of either an inequality of an equality, e.g. 
	$\frac{2y}{10}-1<0$ or $3x+2y - 5=0$. In both cases, an engineer compiles an
	algebraic expression that can be evaluated to produce an output.

	These operations, though very simple, are effectively the same kind of task 
	performed by an analysis block. A constraint or objective function can therfore 
	be represented as another analysis block on the graph, with it's own input and 
	output variable nodes. Although fundamentally no different than an analysis block, 
	for clarity and convience it is useful to distinguish between regular analysis 
	blocks and those that arise from the addition of objectives or constraints to 
	the graph. Therefore, we define an \emph{expression block} as the collection of variable and model 
	nodes related to a given objective or constraint function. Fig.~\ref{f:obj-cons}
	shows a notional problem graph augmented with expression blocks and appropriate 
	connection edges. 
	\begin{figure}[htb!]
	  \begin{center}
		\includegraphics[width=.6\textwidth]{images/obj_const_graph}
	  \end{center}
	  \caption{Notional graph with and objective and constraint nodes \label{f:obj-cons}}
	\end{figure}

\subsection{Local vs Global Nodes}
	\label{ss:local vs global}
	A simple definition of a global node is any node that directly involves information 
	from more than one analysis block. Likewise, a local node is any node 
	that involves information from only a single analysis block. As stated before, variable 
	nodes with fixed edges directed to model nodes (as part of an analysis block) are inherently local 
	and similarly model nodes are also local. Expression and driver nodes
	are not inherently local or global though. For these, their locality is determined 
	by the incomming and outgoing edges they possess. Consider the graph in Fig. \ref{f:local_global}, 
	where the expression node $g_1$ represents some kind of constraint for the optimziation problem. 
	Since $g_1$ only has edges comming from analysis block $B$, it would be considered a local constraint 
	on block $B$. $g_2$ has incomming edges from both $A$ and $B$ analysis blocks, so it would be a global constraint. 
	\begin{figure}
	  \begin{center}
	  %\includegraphics[height=.25\textheight]{XDSM/simple}
	  [[insert graph here]]
	  \caption{A simple graph with local and global nodes \label{f:local_global}}
	  \end{center}
	\end{figure}

@@

\subsection{Coupling Between Analyses}
	\label{ss:coupling}
	Coupling exists in a design problem when a set of two or more analyses each depend on the
	others outputs. Consider the notional problem from Sec.~\ref{s:specific vs fundamental} where two 
	analyses, $A$ and $B$ are coupled through their respective variables. The graph of this 
	is shown in Fig.~\ref{f:coupling}. In the graph $A.z$ has an outgoing connection 
	edge to $B.z$, indicating that there is a connection between the two. 
	Similarly a connection exists between $B.y$ and $A.y$. These two connections form 
	a cycle which represent the coupling between them. This cycle means that blocks $A$ and $B$ need
	to be iterated on until they converge. 
	\begin{figure}
	  \begin{center}
	  %\includegraphics[height=.25\textheight]{}
	  [INSERT GRAPH HERE]
	  \caption{Graph of a notional problem with simple coupling \label{f:coupling}}
	  \end{center}
	\end{figure}

@@

	Larger problems can contain more complex cycles in their FPG, indicating more 
	complex coupling between analyses. A cycle can contain more than 
	just two analysis blocks. Multiple independent cycles can exist, indicating 
	independent coupling relationships. Cycles can also overlap, where the same analysis 
	blocks are involved in multiple different coupling cycles. All of these situations
	arise naturally as the size of problems grows, and managing this coupling may
	become difficult. From the perspective of the graph, the problem reduces down to 
	an ordering challenge. For each cycle in the graph, an analysis block needs to be selected 
	to start the iteration from. Rogers' DEMAID tool uses a genetic algorithm to find an ordering 
	that minimizes the overall coupling of the system by separating independent 
	cycles in the graph \cite{rogers1996,rogers1996demaid}. In the present work, Sec.~\ref{s:example problem}.\ref{ss:obtaining an FPG} will suggest that an FPG may be obtained with respect to the number of cycles or the length of these cycles, etc.

\subsection{Multi--fidelity Problems}
	\label{ss:multi-fideliy problems}
	A multi--fidelity problem can be characterized by two or more different analyses 
	each calculating the same data. Mapping this type of multi--fidelity situation 
	onto a graph will yield two analysis blocks, each with their own output 
	variable nodes, having outgoing edges to a third analysis block that needs the 
	data as input. Figure \ref{f:collision-example} shows a notional problem 
	that exhibits this structure, where the key feature is that variable node $y$ in 
	analysis block $C$ has two incoming edges. 
	\begin{figure}
	  \begin{center}
	  %\includegraphics[height=.25\textheight]{}
	  [INSERT GRAPH HERE]
	  \caption{Graph of a notional multi-fidelity problem \label{f:collision-example}}
	  \end{center}
	\end{figure}

@@

\section{Building MCG and FPG}
	\label{s:building graphs}
\subsection{Maximal Connectivity Graph}
	\label{ss:MCG}
	To construct the maximal connectivity graph, we assume that a set of codes, global inputs, and objectives and constraints (collectively called global outputs) are provided. The codes are represented by analysis block graphs ((ABGs?)) $A_i=(V_{A_i},E_{A_i}), \ i\in I$, where $m$ is the number of codes and $I=\{1,2,\ldots,m\}$. The global inputs are represented as a set of variable nodes $V_\txt{in}$, and the global outputs are represented by a set of variable nodes $V_\txt{out}$. 
	We assume that $V_\txt{in}$, $V_\txt{out}$, and each $A_i$ are given, and that any potential connection between variables is given in the form of connection edges in the set $C_M$. 
	Then we may construct the maximal connectivity graph $M=(V_M,E_M)$ as
	\begin{IEEEeqnarray*}{rCl}
	V_M & = & V_\txt{in} \cup V_\txt{out} \cup \left( \bigcup_{i \in I} V_{A_i} \right), \\
	E_M & = & C_M \cup \left( \bigcup_{i \in I} E_{A_i} \right),
	\end{IEEEeqnarray*}
	The MCG $M$ is uniquely determined by the given set of analysis blocks, the required outputs, and the given global inputs. In the cases where the set of global inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

\subsection{Fundamental Problem Graph}
	\label{ss:FPG}
	We now define the fundamental problem graph $F=(V_F,E_F)$ as a directed graph meeting the following conditions, which are explained subsequently,
	\begin{enumerate}
	\item[(1)] $\displaystyle{I_F \subset I}$ such that the remaining conditions hold
	\item[(2)] $\displaystyle{ \forall i \in I_F, \ \exists v \in V_{A_i} \txt{ with } t_\txt{node}(E^{-1}(v))=\txt{`model'} \txt{ and } \txt{deg}^+(v)>0 }$
	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_{A_i} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	%\item[(3)] $\displaystyle{ \forall i \in I, \forall v \in V_\txt{out} \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) < \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	\item[(3)] $\displaystyle{V_{\txt{in},F} = \{v \in V_\txt{in} \st \txt{deg}^+(v) > 0\} }$ ((maybe this should be 1 instead of 0 to match the previous definition of global input))
	\item[(4)] $\displaystyle{V_F = V_{\txt{in},F} \cup V_\txt{out} \cup \left( \bigcup_{i \in I_F} V_{A_i} \right)}$
	\item[(5)] $\displaystyle{E_F = C_F \cup \left( \bigcup_{i \in I_F} E_{A_i} \right), \ C_F \subset C_M}$
	\item[(6)] $\displaystyle{\forall v \in V_F \txt{ with } t_\txt{node}(v) = \txt{`variable,'} \  \txt{deg}_l^-(v) \leq \txt{deg}^-(v) \leq \txt{deg}_u^-(v)}$
	\end{enumerate}
	The set $I_F$ is an index set containing the indices of the analysis blocks in $F$ and it may only include the analysis blocks $A_i$ that meet requirements (2) and (6). 
	Requirement (2) stipulates that each analysis blocks in $F$ must have at least one local output that is being used. 
	Requirement (3) stipulates that only the global inputs that are being used should be included. 
	Requirements (4) and (5) provide the construction of the sets of nodes and edges, respectively. 
	Lastly, requirement (6) stipulates that the number of edges directed into the variable nodes must be within the lower and upper in-degree limits; if $\txt{deg}^-(v) < \txt{deg}_l^-(v)$ the node is called a \emph{hole}, and if $\txt{deg}^-(v) > \txt{deg}_u^-(v)$ the node is called a \emph{collision}.

@@

\subsection{Obtaining the Fundamental Problem Graph}
	\label{ss:obtaining FPG}
	In general, there may be multiple different graphs that satisfy the FPG conditions, though there may be none at all. Here, we describe a process for obtaining an FPG by starting with the MCG and disconnecting connection edges until the FPG conditions are met. Then the problem is reduced to deciding which connection edges to remove.

### 2013-04-10 intro.tex
cross referencing

\section{Introduction}

    The number of analysis tools required in multidisciplinary design optimization (MDO) studies is growing
    in parallel with the increasing scope of typical MDO problems. An example of this growth in scope can
    be observed in the historical evolution of the disciplines involved in MDO problems in aircraft design.
    Multidisciplinary optimization emerged as a separate field from structural optimization through the
    need to introduce formal techniques for managing the coupling of aerodynamic loads and structural
    deformations, through the linking of aerodynamic vortex lattice or panel methods with structural finite
    element models [[REF]]. Subsequently, flight performance and life cycle economics tools were integrated
    into MDO analysis workflows for conceptual and preliminary design studies [[REF]]. Currently, MDO
    problems for aircraft design also often include tools for aircraft noise and emissions. In sum, it is now
    commonplace for 5-10 analysis tools to be employed for typical aircraft design optimization studies [[REF]].

@@

\section{Specific vs Fundamental Problem Formulation }

    The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
    of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
    is used to solve the problem. For example, consider the following notional problem:

@@

\section{Existing Graph-Based Syntax}

    As shown above, the mathematical language for specifying problem formulations is very general and can be used both for 
    fundamental and specific problem formulations. Tedford and Martins used the above mathematical syntax to specify the 
    FPF for a set of test problems and also to describe specific formulations for solving them with a 
    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
    problem formulations can all relate back to a common FPG. The challenge with using this 
    traditional mathematical syntax is that it is not easily manipulated or analyzed. 
    A number of graph-based methods have been used successfully to translate the 
    mathematical syntax into a more useful computational form.

>>>>

\section{Introduction}
    \label{s:intro}
    The number of analysis tools required in multidisciplinary design optimization (MDO) studies is growing
    in parallel with the increasing scope of typical MDO problems. An example of this growth in scope can
    be observed in the historical evolution of the disciplines involved in MDO problems in aircraft design.
    Multidisciplinary optimization emerged as a separate field from structural optimization through the
    need to introduce formal techniques for managing the coupling of aerodynamic loads and structural
    deformations, through the linking of aerodynamic vortex lattice or panel methods with structural finite
    element models [[REF]]. Subsequently, flight performance and life cycle economics tools were integrated
    into MDO analysis workflows for conceptual and preliminary design studies [[REF]]. Currently, MDO
    problems for aircraft design also often include tools for aircraft noise and emissions. In sum, it is now
    commonplace for 5-10 analysis tools to be employed for typical aircraft design optimization studies [[REF]].

@@

\section{Specific vs Fundamental Problem Formulation }
	\label{s:specific vs fundamental}
    The Fundamental Problem Formulation (FPF) for any given problem will be constant regardless 
    of which MDAO framework, optimization algorithm, iterative solver, or solution strategy
    is used to solve the problem. For example, consider the following notional problem:

@@

\section{Existing Graph-Based Syntax}
	\label{s:existing syntax}
    As shown above, the mathematical language for specifying problem formulations is very general and can be used both for 
	fundamental and specific problem formulations. Tedford and Martins used the above mathematical syntax to specify the 
	FPF for a set of test problems and also to describe specific formulations for solving them with a 
    number of optimization architectures\cite{Tedford2009}. Their work demonstrates clearly how multiple specific 
    problem formulations can all relate back to a common FPG. The challenge with using this 
    traditional mathematical syntax is that it is not easily manipulated or analyzed. 
    A number of graph-based methods have been used successfully to translate the 
    mathematical syntax into a more useful computational form.
